---
title: "vignette"
author: "Bayesian Econ"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette proivdes a small example that illustrates how to use `MCMC_ADRS_AD` and the other functions that come with the package.

## Circular Example 

We start by generating a network with $N$ nodes from the following latent space specification:

$$y_{ij} \sim Poi(\lambda_{ij}) $$
where:

$$\log \lambda_{ij} = \beta - ||\bf{x}_i-\bf{x}_j||^2$$
In tis example, we will assume that nodes lie on a $K$ dimensional circle.

```{r, fig.show='hold'}
####Libraries#########
require(Rcpp)
require(data.table)
require(AMRSforLS)
#####################

N = 20 #number of nodes
K = 2  #number of latent coordinates

beta  = 5 #intercept
POIS = matrix(0, N, N) #empty matrix

#Setup the latent space representation of nodes on a cirlce

X = matrix(rep(0, N*K), N, K) #X stores the latent coordinates

r <- 1 #radius of the circle

# Calculate the angle between each point
angle <- 2 * pi / N

# Create empty vectors to store the coordinates
x <- numeric(N)
y <- numeric(N)

# Generate the coordinates of each point
for (i in 1:N) {
  x[i] <- r * cos(i * angle)
  y[i] <- r * sin(i * angle)
}

X[,1]<- x
X[,2]<- y

plot(X[,1], X[,2], xlab = "coordinate 1",
     ylab = "coordinate 2", pch = 19, col ="steelblue") #plot the latent positions of the nodes

D = as.matrix(dist(X))^2

for(i in 1:N){
  for(j in (i):N){
    POIS[i,j] = rpois(1, exp(beta -D[i,j]))
  }
}

ADJ = t(POIS) + POIS
diag(ADJ) = 0
rm(D, POIS)
```

## Estimation
We use the function `MCMC_ADRS_AD`, similarly one can use `MCMC_ADRS_BLOCK_AD` and `MCMC_AD`.

```{r, fig.show='hold'}

 N = nrow(ADJ)
  K = 2 #set the dimensions of the latent space

  zi = matrix(0, N , K)  #set empty starting points

  #add noise to starting points
  for(k in 1:K){
    zi[,k] = rnorm(N,0, 0.01)
  }

  #set the adaptive scaling parameters, Andrieu and Thoms 2008
  lam_ad = rep(2.8 , N)
  mu_mat = matrix(0, N, K)
  id<-diag(rep(0.01, K))
  Sigma_ad <- rep(list(id), N)
  Sigma_ad <- lapply(Sigma_ad, FUN = function(x){as.matrix(x)})

  #start the estimation
  result =MCMC_ADRS_AD( y = ADJ, #adj matrix
                         beta = 0,   #intercept starting point
                         beta_a = 0,   #mean prior on beta
                         beta_b = 5,  #sd prior on beta
                         z_i = zi,  # starting points for latent coordinates
                         zeta_a = 0, #mean prior on latent coordinates
                         zeta_b = 5,  #sd prior on latent coordinates
                         lam_ad, #adaptive scaling parameters, see help
                         mu_mat, #adaptive scaling parameters, see help
                         Sigma_ad,#adaptive scaling parameters, see help
                         N = N, #number of nodes
                         K = K,  #dimension latent space
                         prop_sigma_beta  = 0.05, #proposal sd beta
                         acc_target_z  = 0.234, #proposal sd zeta
                         scan_every = 100,
                         Iterations =30000, #number of
                         k_shift = 0.5,
                         eq_option =1, #equal probability option
                         eq_prob= 0.25  #if eq_option =1, equal probability
                         )
```

## Explore the Chains

We now explore the list of results and compare it with the groundtruth.

### Intercept $\beta$
```{r, fig.show='hold'}
plot.ts(result$beta_ite, type = "l", xlab = "iteration", ylab = expression(beta), col = "steelblue")
abline(h = beta, col = "red", lty = 3)
```

### Latent Coordinates $\bf{x}_i$

```{r, fig.show='hold'}
zi_res<-lapply(result$zeta_ite, FUN = function(x){apply(x,2, function(x){x-mean(x)})}) #de-mean
zi_res<-lapply(zi_res, FUN = function(x){procrustes_cpp(as.matrix(X[,1:K]),x )}) #apply procrustes trasnformation
 zi_sub<-zi_res[seq(2000,3000, 10)] #burnin and thinning
 
 zi_sub<-lapply(zi_sub, function(x){data.table::data.table(x)})
 Zall<-data.table::rbindlist(zi_sub)
 Zall<-data.frame(Zall)

 #plot the estimated latent coordinates  
 plot(Zall$V1, Zall$V2, xlab = "coordinate 1",
     ylab = "coordinate 2", pch = ".", col ="steelblue") 
 #plot the groundtruth in red
 points(X[,1], X[,2], col = "red", pch = 4)
```

